---
title: "STAT600 - Automated EDA in R"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

After getting some new dataset, doing a proper exploratory data analysis (EDA) is very important part. This help to better undestand the structure of our data, potential problems and project the steps needed to use the particular data set for modelling. 

Most of the data analysts would agree that data preparation (cleaning, putting to the correct format) and exploration of a data set takes big chunk of the time devoted to the project. 

(Un)Fortunately people are lazy, thus some started developing some ways how to spead up the EDA part. 

Here I present 4 R libraries that has some functions for automatic EDA - `DataExplorer`, `Hmisc`, `dlookr`, `explore`.

```{r}
library("DataExplorer")
library("Hmisc")
library("dlookr")
library("exploreR")
```

For each of the package, the data used comes from suggestions of packages itself. 

##DataExplorer

This package is created by Boxuan Cui. Here are showed three functions from DataExplorer package - `introduce()`, `plot_intro()`, and `create_report()`. The datasets used for a presentation - airquality and diamonds. 

```{r}
DataExplorer::introduce(airquality)
```

```{r,warning = FALSE, results='hide',message=FALSE}
DataExplorer::plot_intro(airquality)
```

```{r, warning = FALSE, results='hide',message=FALSE}
DataExplorer::create_report(diamonds, y="price")
```



If you interested in more packages and their comparison, there are some references:
 - The Landscape of R Packages for Automated Exploratory Data Analysis by Mateusz Staniak and Przemysław Biecek
 - https://github.com/mstaniak/autoEDA-resources
