---
title: "STAT600 - Automated EDA in R"
output:
  output: rmarkdown::github_document
---


## Introduction

After acquiring some new data, doing a proper exploratory data analysis (EDA) is a very important part. This help to better understand the structure of our data, potential problems and project the steps needed to use the particular data set for modeling. 

Most of the data analysts would agree that data preparation (cleaning, putting to the correct format) and exploration of a data set takes a big chunk of the time devoted to the project. 

(Un)Fortunately, people are lazy, thus some started developing some ways how to speed up the EDA part. 

Today I will present three `R` libraries for automatic EDA - `DataExplorer`, `Hmisc`, `dlookr`.

Notice that there are way more libraries in the field and each is doing some different. References for more libraries are given in the end. 

```{r,warning = FALSE, results='hide',message=FALSE}
library("DataExplorer")
library("Hmisc")
library("dlookr")
```

For presentation of function I used two very popular data sets - airquality and diamonds. 


## DataExplorer

This package is created by Boxuan Cui. Below are presented three functions from DataExplorer package - `introduce()`, `plot_intro()`, and `create_report()`.  

```{r}
DataExplorer::introduce(airquality)
```

```{r,warning = FALSE, results='hide',message=FALSE}
DataExplorer::plot_intro(airquality)
```

```{r, warning = FALSE, results='hide',message=FALSE}
DataExplorer::create_report(diamonds, y="price", output_file="DataExplorer_diamonds_report.html")
```

Notice, that `create_report()` generates the separate file `DataExplorer_diamonds_report.html`. 

## Hmisc

This package is created by Frank E Harrell Jr, with contributions from others. Below is showed one function from `Hmisc` package - `describe()`.

```{r}
Hmisc::describe(airquality)
```

Similar results with some additional information could be presented as more user-friendly version. 

```{r, warning = FALSE, results='hide',message=FALSE}
p <- Hmisc::describe(airquality)
pp <- Hmisc::html(p)
htmltools::save_html(pp, file="Hmisc_describe_airquality.html")
```

Results are saved into the file `Hmisc_describe_airquality.html`. 

## dlookr

This package belongs to the tidyverse universe. Here are presented function `eda_report()`.


```{r,warning = FALSE, results='hide',message=FALSE}
dlookr::eda_report(diamonds, output_format="html",
                   target="price", 
                   output_file="dlookr_eda_report_diamonds.html",
                   output_dir=getwd())
```

Results are saved into the file `Hmisc_describe_airquality.html`. 


## Wrap up

If you interested in more packages and their comparison, there are some references:
 * Overview - The Landscape of R Packages for Automated Exploratory Data Analysis by Mateusz Staniak and Przemysław Biecek. M
 * Collection of smart EDA packages: https://github.com/mstaniak/autoEDA-resources
